### 2014-03-17 paper/results.tex

\section{MRI in self-gravitating disks}
In this section we focus on the MRI. We use the vertical field setup
of case 1. We first consider MRI modes with negligble densit/potential
perturbations to see the effect of self-gravity on the MRI through the
background stratification, then go on to consider MRI modes with
significant density/potential perturbations in massive disks.

\subsection{Influence of self-gravity on the MRI through the
  background equilibrium}

Here we use polytropic disks, which have a well-defined
thickness. The upper disk boundary is set to 
$Z_s=0.99H$. We fix $\beta=100$ and $k_xH=0.1$ unless otherwise
stated.

%kx -> 0 so incompressible

\subsubsection{Uniform
  resistivity}  
We first calculate the MRI as a function of $Q$ with uniform
resistivity ($A=1$) for mid-plane Elsasser numbers $\Lambda_0\in[0.3,10]$.

Fig. \ref{compare_growth_poly_uniresis} plots the MRI growth rates as
a function of $Q$ and $\Lambda_0$. As expected growth rates generally
decrease with increasing $\Lambda_0$. In the limit of ideal MHD ($\Lambda_0>1$), 
there is negligible dependence on $Q$. In the resistive limit
($\Lambda_0<1$), however, growth rates decrease noticeably for $Q<1$. This shows
show that disk self-gravity affects the MRI through the background
equilibrium, even though density and potential perturbations are  
negligible (i.e. the linear response is
non-self-gravitating).

\begin{figure}
  \includegraphics[width=\linewidth]{figures/compare_growth_poly_uniresis2}
  \caption{MRI growth rates as a function of $Q$ for mid-planet
    Elsasser numbers $\Lambda_0=0.3$ (solid), $0.5$ (dotted), $1.0$
    (dashed) and $5.0$ (dash-dot).  
    \label{compare_growth_poly_uniresis}}
\end{figure}

\cite{sano99} found that for instability to occur the linear mode
wavelength $\lambda$ should fit inside the disk. That is,   
\begin{align}\label{sano_crit}
  \lambda \equiv
  \mathrm{max}\left(\lambda_\mathrm{ideal},\lambda_\mathrm{resis}\right)\lesssim
  2H, 
\end{align}
where the MRI wavelengths are given by 
\begin{align}\label{lambda_ideal}
  \frac{\lambda_\mathrm{ideal}}{2H} = \frac{4\pi}{\sqrt{15}} f \hat{v}_A =
  \frac{4\pi f}{\sqrt{15\beta\hat{\rho}}}
\end{align}
in the limit of ideal MHD, and 
\begin{align}\label{lambda_resis}
  \frac{\lambda_\mathrm{resis}}{2H} = \frac{2\pi}{\sqrt{3}}\frac{\hat{\eta}}{\hat{v}_A f} =
  \frac{2\pi f}{\Lambda_0}\sqrt{\frac{\hat{\rho}}{3\beta}} 
\end{align}
in the limit of high resistivity.

Since $\hat{\rho}$ is weakly dependent
on $Q$ (Fig. \ref{eqm_den}), self-gravity only affects the
MRI through the factor $f$, which increases with decreasing $Q$ (see
Fig. \ref{plot_fq} in Appendix \ref{appen1}). This suggests that
sufficiently strong self-gravity can stabilize the MRI by making
$ 2H<\lambda $.

In the ideal limit, we find $\lambda < 2H$ throughout most of the disk
for the values of $Q$ considered, so that self-gravity does not affect
growth rates significantly. However, the wavelength of the instability
increases. This is shown in Fig. \ref{compare_result_lambda10} which
plots the magnetic energies for $\Lambda_0=10$ and a range of Toomre $Q$
values. The number of vertical nodes decrease with $Q$, 
i.e. the disk accommodates fewer wavelengths because increasing
vertical self-gravity makes it thinner.

%Weak variations in growth rates in the ideal limit, seen in 
%Fig. \ref{compare_growth_poly_uniresis}, were found to be associated
%with changes in the mode character: $\gamma$ decreases
%slightly when the mode begins to acquire an additional vertical node as $Q$ is
%increased.

\begin{figure}
  \includegraphics[width=\linewidth]{figures/compare_result_lambda10}
  \caption{Magnetic energies as a function of height, in the limit of
    ideal MHD, for various strengths of self-gravity.   
    \label{compare_result_lambda10}}
\end{figure}

Self-gravity appreciably decreases the MRI growth rates in the
resistive limit. Fig. \ref{lambda_poly_resis} plots 
Eq. \ref{sano_crit} for $\Lambda_0=0.3$. In the non-self-gravitating
disk ($Q=4$) the instability criterion is marginally satisfied and the
MRI operates. As $Q$ decreases, 
Eq. \ref{sano_crit} is violated and the MRI growth rate is
significantly reduced. This is seen for $Q=0.2$ where $\lambda \geq 2H$ throughout
the disk. (The instability is not completely surpressed since
Eq. \ref{sano_crit} is only an approximation.) Although the function
$f$ does not change significantly for the range of $Q$ considered, the
dependence of $\lambda_\mathrm{resis}$ on $f$ (and therefore $Q$) is
amplified by the denominator $\Lambda_0<1$ in the resistive case. For
resistive cases we find no nodes in the magnetic energy $E_1$ except
at end points, i.e. only the longest wavelength survives.

\begin{figure}
  \includegraphics[width=\linewidth]{figures/lambda_poly_uniresis}
  \caption{Approximate wavelengths of the most unstable resistive MRI modes as given by
    Eq. \ref{sano_crit}---\ref{lambda_resis}, normalized by the 
    disk thickness, as a function of height for a range of Toomre $Q$
    values.  
    \label{lambda_poly_resis}}
\end{figure}

\subsubsection{Layered
  resistivity} 
Here we consider disks with mid-plane Elsasser number $\Lambda_0=0.1$
and a variable resistivity profile with
$A=10^2$. Fig. \ref{poly_layer} compares the magnetic  
energies for $Q=0.2,\,1$ and $4$. They have similar growth rates, $\gamma/\Omega
= 0.53,\,0.64$ and $0.66$, respectively. In the non-self-gravitating
limit ($Q=4$), the MRI is effectively surpressed for
$z\lesssim0.5H$. This is consistent with layered accretion. However,
in the massive disk ($Q=0.2$) the mode occupies a wider vertical
extent because its wavelength (in units of $H$) is larger. This
suggests that in massive disks the MRI is not well localized to the
active layer.

%MRI penetrates 
%d is not as well localized

\begin{figure}
  \includegraphics[width=\linewidth]{figures/compare_results_poly_layer_amp100}
  \caption{Magnetic energies as a function of height, for layered polytropic disks with
    such that the conductivity increases by a 
    factor of $10^2$ in going from the mid-plane to the upper disk
    boundary. The vertical lines indicate $\Lambda=1$ for each value
    of $Q$.
    \label{poly_layer}}
\end{figure}

\subsubsection{Dependence on $k_x$}
The above experiments show that with increasing 
disk self-gravity, the MRI becomes more global in the vertical
direction. We find a similar result in the horizontal direction. 
In Fig. \ref{compare_growth_poly_kx} we plot MRI growth rates as a
function of $k_x$ for a range of $Q$ values. Increasing self-gravity
decreases the cut-off horizontal wavenumber for the MRI, implying that
small-scale MRI does not operate in massive disks.

\begin{figure}
  \includegraphics[width=\linewidth]{figures/compare_growth_poly_varQ_kx}
  \caption{MRI growth rates in self-gravitating polytropic disks, as a
    function of the horizontal wavenumber $k_x$. The disk is ideal
    ($\Lambda_0=10^2,\, A=1$) with $\beta = 40$. These modes have negligible
    density/potential perturbations.  
    \label{compare_growth_poly_kx}}
\end{figure}

%these modes are still non-sg
%use beta = 40 so only a small range of kx need to be simulated to
%demonstrate this effect
%{\bf cut-off $k_xH$ decreases with increasing SG -> small-scale MRI
%  stabilized by self-gravity}

\subsection{Influence of self-gravity on the MRI through the linear
  response}  
Our goal here is to examine whether or not self-gravity can amplify
the density perturbations caused by the MRI. We choose a massive 
isothermal disk with $Q=0.2$ (corresponding to $Q_\mathrm{2D}=0.72$),
which is still expected to be marginally 
stable to gravitational instability \citep[][who find
a critical value of  $Q\simeq 0.2$]{mamat10}.   
The upper disk boundary is set to $Z_s=H$.

%resolution Nz=256

\subsubsection{Ideal case}
We first consider the limit of ideal MHD by adopting a uniform
resistivity with $\Lambda_0=100$. %Non-zero density perturbations
%require $k_x\neq0$ so we consider linear modes as a function of $k_x$.   
Fig. \ref{gravity_energy} shows the mean gravitational potential
energy perturbation, as a function of $k_x$ for several values of
$\beta$. The potential perturbation is negligible for all cases when 
$k_xH\lesssim 0.5$, since the MRI becomes incompressible as $k_x\to
0$.

\begin{figure}
  \includegraphics[width=\linewidth]{figures/compare_energy_ideal}
  \caption{Average gravitational potential energy perturbation
    associated with the MRI, compared to the average magnetic energy
    perturbation, in a self-gravitating disk with $Q=0.2$ in 
    the limit of ideal MHD ($\Lambda_0=10^2$), for a range of field
    strengths $\beta$.  
    \label{gravity_energy}}
\end{figure}

For $\beta\gg 1$, i.e. a weak field, density perturbations are
negligible and the standard MRI operates. However, as $\beta$ is
lowered we find non-negligible potential perturbation for
$k_xH=O(1)$. This suggets that in a strongly magnetized disk that
still permits the MRI, the associated density perturbation can be
important when the disk is self-gravitating.

Fig. \ref{mri_massive} shows an example of a MRI mode with
non-negligible density perturbation. The growth rate $\gamma\simeq
0.4\Omega$ is comparable to that of the most unstable mode
($\gamma\simeq 0.6\Omega$ for $k_x\simeq0$). 
We therefore expect significant density perturbations to grow
dynamically, even though the disk is expected to be gravitationally
stable.

%dynamical
%timescales 
%($\gamma\sim0.4\Omega$)

\begin{figure}
  \includegraphics[width=\linewidth]{figures/result_ideal_sg}
  \caption{Example of a linear MRI mode with non-negligible density and
    potential perturbation, in a strongly magnetized massive disk
    ($\beta=10$ and $Q=0.2$) in the limit of ideal MHD 
    ($\Lambda_0=10^2$).   
    \label{mri_massive}}
\end{figure}

\subsubsection{Resistive case}
We repeat the above calculation for resistive disks, but fix
$\beta=100$. Fig. \ref{gravity_energy_resis} shows the 
perturbed gravitational energy increases with   
resistivity. Interestingly, in the highly resistive case
the MRI can be energeically dominated by self-gravity, i.e. the case
with $\Lambda_0=0.1$ and $k_xH\simeq1.3$. We also calculated this mode
in the \emph{Cowling  
  approximation}, where the Poisson equation is ignored in the 
linearized equations and the potential perturbation set to zero  
(formally letting $Q\to\infty$ in Eq. \ref{lin_poisson}).
Fig. \ref{mri_massive_cowling} compares the magnetic energy in the
Cowling approximation to that of 
the full problem in . The growth rate increases when self-gravity is
included in the linear response, as expected, but only
slightly. $\gamma$ and $E_m$ are similar in both cases, indicating
that the instability in the self-gravitating calculation is 
still the MRI.

\begin{figure}
  \includegraphics[width=\linewidth]{figures/compare_energy_resis}
  \caption{Average gravitational potential energy perturbation
    associated with the MRI, compared to the average magnetic energy
    perturbation, in a self-gravitating disk with $Q=0.2$,
    for a range of midplane Elsasser numbers $\Lambda_0$ at fixed
    field strength $\beta=100$.    
    \label{gravity_energy_resis}}
\end{figure}

\begin{figure}
  \includegraphics[width=\linewidth]{figures/compare_result_cowling}
  \caption{Magnetic energy associated with the linear mode with
    largest gravitational-to-magnetic energy ratio in
    Fig. \ref{gravity_energy_resis} (solid) compared with that computed
    under the Cowling approximation (dotted). 
    \label{mri_massive_cowling}}
\end{figure}

%An example is shown in
%Fig. \ref{mri_massive_resis}, which  plots the energies assocaited with the 
%linear MRI mode with $k_xH\sim1.3$ in the $\Lambda_0=0.1$  
%disk. The gravitational potential energy significantly exceeds the
%magnetic energy near the midplane ($z\lesssim0.3H$).

%\begin{figure}
%  \includegraphics[width=\linewidth]{figures/result_resis_sg}
%  \caption{Example of a linear resistive MRI mode with significant
%    gravitational potential energy perturbation.  The vertical line
%    inidicates $\Lambda=1$.  
%    \label{mri_massive_resis}}
%\end{figure}

\subsubsection{Qualitative interpretation} 
In order to make sense of the above results, we return to ideal MHD. 
We consider regions close to the disk midplane ($z\sim 0$), where
self-gravity is expected to be most important. For this discussion we
will ignore stratification and set $d/dz\to \imgi k_z$ where $k_z$ is
a vertical wavenumber. The governing equations are then

\begin{align}
  &  0= v_A^2k^2\dvx + \imgi\sigma\left(\imgi\sigma \dvx - 2\Omega\dvy + \imgi k_x \w\right),\label{simp1}\\
  &  0= v_A^2k_z^2\left(\dvy + \frac{\imgi S
  }{\sigma}\dvx\right) + \imgi\sigma\left(\imgi\sigma\dvy +
  \frac{\kappa^2}{2\Omega}\dvx\right),\label{simp2}\\
  & 0 = -k_z^2\w + \frac{\sigma^2}{c_s^2}W + \sigma k_x \dvx, \label{simp3}\\
  & 0 = k^2\dphi + \frac{\Omega^2}{c_s^2Q}W,\label{simp4}
\end{align}
where $k^2 = k_z^2 + k_x^2$.

We imagine an iterative procedure to solve the above equations,
starting from the Cowling approximation. %defined to be the
%solution to Eq. \ref{simp1}---\ref{simp3} with potential perturbation
%set to zero. 
This is the standard MRI and we denote the solution as
$\dvx^{(0)}$, $\dvy^{(0)}$ and $W^{(0)}$.

We now include self-gravity. In general, the compressible MRI has
$W^{(0)}\neq0$. The Poisson equation implies
an associated potential perturbation,    
\begin{align} 
  \dphi = -\frac{\Omega^2}{c_s^2 Q k^2} W^{(0)}.
\end{align}
Physically, we expect $k^2\geq0$, so that a positive (negative) density
perturbation causes a negative (positive) potential perturbation. We
then regard $\dphi$ to alter the Cowling solution, so
that $\dvx^{(0)} \to \dvx^{(0)} + \dvx^{(1)}$ and similarly for $\dvy$ and
$W$. Then 
\begin{align}
  &   k_x\sigma \dphi = v_A^2k^2\dvx^{(1)} + \imgi\sigma\left[\imgi\sigma
  \dvx^{(1)} - 2\Omega\dvy^{(1)} + \imgi k_x W^{(1)}\right], \label{simp_pert1}\\ 
  &  0= v_A^2k_z^2\left[\dvy^{(1)} + \frac{\imgi S
    }{\sigma}\dvx^{(1)}\right] + \imgi\sigma\left[\imgi\sigma\dvy^{(1)} +
  \frac{\kappa^2}{2\Omega}\dvx^{(1)}\right],\label{simp_pert2}\\
  & k_z^2\dphi  = \left(\frac{\sigma^2}{c_s^2}-k_z^2\right)W^{(1)} +
  \sigma k_x \dvx^{(1)}.\label{simp_pert3} 
\end{align}
Now, if the perturbations to the magnetic field remain
unchanged, i.e. the mode remains close to the standard MRI, then
$\dvx^{(1)} \sim 0$ and $\dvy^{(1)}\sim0$, so Eq. \ref{simp_pert2} is
satisfied. Eq. \ref{simp_pert1} then require $\dphi +
W^{(1)}\sim0$. This is compatible with Eq. \ref{simp_pert3} if 
\begin{align}
  \left|k_z^2\right| \gg \left|\frac{\sigma^2}{c_s^2}\right|. \label{cond}
\end{align}
For the ideal MRI,  we assume $k_z\sim \Omega/v_A$. Then
$|\sigma^2/c_s^2k_z^2|\sim |\sigma^2/\Omega^2\beta|\ll1$ because
$|\sigma|\lesssim \Omega$ and we are considering $\beta\gtrsim
10$. Thus Eq. \ref{cond} is generally satisfied.

The above assumptions imply
\begin{align}
  W^{(1)} \sim \frac{\Omega^2}{c_s^2 Q k^2} W^{(0)},\label{feedback}
\end{align} 
which indicates a non-zero density perturbation due to the
MRI can be amplified by self-gravity. Now, for $k_xH\sim 1$ we have
$|k_z^2/k_x^2|\sim \beta/f\gg1$ because $f=O(1)$ and $\beta\gtrsim10$
for the cases considered above. Then
\begin{align}
  \left|\frac{W^{(1)}}{W^{(0)}}\right| \sim \frac{1}{Q\beta}, 
\end{align}
 suggesting stronger amplification of the density field by 
 self-gravity with increasing field strength (decreasing
 $\beta$).

 For weak fields, the MRI has small vertical wavelength,
 implying rapid density variations with height. The magnitude of the
 corresponding gravitational potential is then expected to be small
 because of internal cancellations, since the potential is essentially
 the integral of the density field.

% not expect significant potential perturbations

%Conversely, for strong fields, only a single MRI wavelength occupies
%the fluid column.

The above arguments also apply to the resistive
limit. Eq. \ref{simp3}---\ref{simp4} are unchanged, while resistive
terms appearing in Eq. \ref{simp1}---\ref{simp2} only involve the
potential perturbation through $\w$. For the resistive MRI we assume
$k_z\sim v_A/\eta$ and $|\sigma|\sim v_A^2/\eta = \Lambda\Omega$
\citep{sano99}. Then $|\sigma^2/c_s^2k_z^2|\sim 1/\beta \ll 1$ so
Eq. \ref{cond} is satisfied. Noting that $k_z^2\sim
\Lambda^2\Omega^2\beta/c_s^2$, the feedback equation becomes
\begin{align}
  \left|\frac{W^{(1)}}{W^{(0)}}\right| \sim
  \frac{1}{Q\left(f^2\hat{k}_x^2 + \beta\Lambda^2\right)},
\end{align}
so increasing the resistivity (decreasing $\Lambda$) should enhance
density perturbations.