### 2014-03-20 paper/results_azi.tex
tex update

\section{Effect of an azimuthal field}
In this section we add an azimuthal field so that $B_y\neq 0$,  
parameterized by $\epsilon \equiv B_y/B_z$. Note that we continute to
use $B_z$ for normalizations and $\beta$ is associated with the
vertical Alfven speed. We also extend the previous calculations to the
full disk $z\in[-Z_s,Z_s]$, which allows us to compare the effect of
self-gravity on MRI modes with different symmetries across the
midplane.

\subsection{Fiducial case}
Here we consider an isothermal disk with $Q=0.2$
($Q_\mathrm{2D}=0.72$) and $\beta=10$ in the limit of ideal MHD
($\Lambda_0=100$). We do not expect gravitational instability, since
Fig. \ref{compare_growth3} shows that even for $Q=0.18$, GI is
surpressed for $\beta \lesssim 15$. We consider azimuthal field
strengths with $\epsilon=[0,3]$.

Fig. shows the MRI growth rates for the above system as a function of
$k_x$. The top panel are modes with even midplane symmetry
$W^\prime(0)\sim 0$ for $k_x\to0$, and the bottom panel are modes with odd midplane 
symmetry $W(0)\sim 0$ for $k_x\to0$. The latter set of modes were
excluded in the previous sections by midplane boundary
conditions. The plot also include growth rates computed in the Cowling
approximation. As expected, none of the modes are energetically
dominated by self-gravity.

%general stabilization of MRI

\begin{figure}
  \includegraphics[width=\linewidth,clip=true,trim=0cm 2cm 0cm 0cm]{figures/compare_growth3_tilted_even.ps} 
  \includegraphics[width=\linewidth,clip=true,trim=0cm 0cm 0cm 0.52cm]{figures/compare_growth3_tilted_odd.ps}
  \caption{Growth rates in massive isothermal disks with $Q=~0.2$ and
    $\beta=10$ for a range of azimuthal field strengths $B_y/B_z$. The
    dots are solutions computed from the full problem, with the
    colorbar measuring the gravitational potential perturbation, while
    the solid curves are computed from the Cowling approximation. The
    topic panel correspond to even modes and the
    bottom panel correspond to odd modes, which have
    $W^\prime(0)\sim0$ and $W(0)\sim0$, respectively, for $k_x\to0$.
    \label{compare_growth3_tilted}}
\end{figure}

%We consider $\epsilon\equiv B_y/B_z \in[0,3]$. The

Consider first the even modes in the top panel of
Fig. \ref{compare_growth3_tilted}. Self-gravity increasingly stabilizes
modes with small $k_x$ as $B_y$ is increased, but the effect is
negligible. SG destabilizes modes with  $k_xH\gtrsim
O(1)$. Consequently the cut-off wavenumber is larger when SG is
included. The destabilization effect weakens with $B_y$ because the
total magnetic pressure increases. %which generally opposes GI 
The Cowling approximation becomes 
increasingly good for even modes as $B_y$ is increased.

%cowling is good with increasing azi field for even modes. 
%cannot consider superthermal fields
%compressibility important f

The odd modes in the bottom panel of Fig. \ref{compare_growth3_tilted}
display opposite behavior. Self-gravity has no effect when
the field is purely vertical ($B_y=0$). SG stabilizes all wavelengths,
and is most effective with increasing $B_y$ and $k_xH =  O(1)$. 
%We therefore expect odd modes to become less important as the azimuthal
%field strength is increased.

%compare eigenfunctions

With $B_y\neq 0$, the MRI is compressible even in the Cowling
approximation \citep{pessah05}.

\begin{figure}
  \includegraphics[width=\linewidth,clip=true,trim=0cm 1.5cm 0cm 0cm]{figures/result_tilted_kx0d1.ps} 
  \includegraphics[width=\linewidth,clip=true,trim=0cm 0cm 0cm 0.cm]{figures/result_tilted_kx1d2.ps}
  \caption{caption
    \label{result_tilted}}
\end{figure}

\subsection{Resistive disks}

\subsection{Disks with GI}