### 2014-03-27 paper/paper.tex
tex updates

%\documentclass[12pt]{aastex}
\documentclass[iop]{emulateapj}
%\documentclass{article}
%\usepackage{draftwatermark}
\usepackage{paralist}
\shorttitle{Magnetized massive disks}
\shortauthors{M-K. Lin}

\usepackage{amsmath}
%\usepackage{url}
\usepackage{bm}
%\usepackage[font=small,format=plain,labelfont=bf,up,textfont=it,up]{caption}
\newcommand{\p}{\partial}
\newcommand{\lmax}{l_\mathrm{max}}
\newcommand{\sbar}{\bar{\sigma}} 
\newcommand{\imgi}{\mathrm{i}} 
\newcommand{\ii}{\mathrm{i}}
\newcommand{\C}{\mathcal{C}^\lambda}
\newcommand{\wop}[1]{\mathcal{V}^{(#1)}}
\newcommand{\avg}[1]{\langle{#1}\rangle}
\newcommand{\zhat}{\hat{z}}
\newcommand{\rhat}{\hat{r}}
\newcommand{\he}{\mathcal{H}}
\newcommand{\amp}{\mathcal{A}}
\newcommand{\dd}{\delta}
\newcommand{\real}{\operatorname{Re}}
\newcommand{\imag}{\operatorname{Im}}
\newcommand{\sgn}{\operatorname{sgn}}
\newcommand{\zmax}{Z_s}
\newcommand{\wbar}{\tilde{W}}
\newcommand{\qbar}{\tilde{Q}}
\newcommand{\dbigH}{\frac{H^\prime}{H}}
\newcommand{\Rp}{\frac{\rho_0^\prime}{\rho_0}}
\newcommand{\Fp}{\frac{g^\prime}{g}}
\newcommand{\Rpp}{\frac{\rho_0^{\prime\prime}}{\rho_0}}
\newcommand{\Fpp}{\frac{g^{\prime\prime}}{g}}
\newcommand{\ils}{L_s^{-1}}
\newcommand{\ilp}{L_p^{-1}}
\newcommand{\ihp}{H_p^{-1}}
\newcommand{\ihs}{H_s^{-1}}
\newcommand{\ciso}{c_\mathrm{iso}}
\newcommand{\csmid}{c_{s0}} 
\newcommand{\hiso}{h_\mathrm{iso}}
\newcommand{\Hiso}{H_\mathrm{iso}}
\newcommand{\zeus}{\texttt{ZEUS-MP }}
\newcommand{\rsph}{r_\mathrm{sph}}
\newcommand{\dvx}{\dd v_x}
\newcommand{\dvy}{\dd v_y}
\newcommand{\dvz}{\dd v_z}
\newcommand{\dbx}{\dd B_x}
\newcommand{\dby}{\dd B_y}
\newcommand{\dbz}{\dd B_z}
\newcommand{\w}{ \widetilde{W}}
%\newcommand{\sbar}{\bar{\sigma}}
\newcommand{\dphi}{\dd \Phi}
\newcommand{\actaa}{Acta Astron.}
\defcitealias{lin12}{L12}

\begin{document}

\title{Linear stability of magnetized massive protoplanetary disks}

\author{Min-Kai Lin }%(+Menou, Fromang, Youdin)}
\affil{Canadian Institute for Theoretical Astrophysics,
  60 St. George Street, Toronto, ON, M5S 3H8, Canada}
\email{mklin924@cita.utoronto.ca}

\begin{abstract}
  Magneto-rotational instability (MRI) and gravitational instability
  (GI) are the two principle routes to turbulent angular momentum
  transport in accretion disks. Protoplanetary disks may develop
  both. The goal of this paper is to reinvigorate interest in the study of
  magnetized  massive protoplanetary disks, starting from the basic
  issue of stability. The linear stability of a local patch of a
  self-gravitating, magnetized three-dimensional (3D) disk to
  axisymmetric perturbations is examined. %Addition effects such as
  %resistivity and azimuthal fields are considered.

%We examine the local linear stability of
%  magnetized, self-gravitating and differentially rotating disks to
%  axisymmetric perturbations. 
%
%not many studies model both
%revisits 
\end{abstract}

\section{Introduction}
Astrophysical disks host a wide range of fluid instabilities. Among
them, the magneto-rotational instability \citep[MRI,][]{chandrasekhar61,
  balbus91,balbus98} %balbus %chandrasekhar ref
and gravitational instability \citep[GI, ][]{toomre64,goldreich65a,goldreich65b} 
provide robust pathways to turbulent angular momentum transport that
enables mass accretion \citep[][ 
  and references therein]{balbus99,armitage11,turner14}.   
%gammie, lin/pringle, paczynski 
%relevance of MRI and GI to planet formation 
They are also relevant to planet formation theory. For example,
the strength of MRI turbulence directly affect planetesimal dynamics
in protoplanetary disks \citep{yang12,gressel12}; while GI can
potentially form giant planets directly through disk fragmentation
\citep{boss97,boss98,gammie01,voro13,helled14}. 
%gammie,
%boss %some review articles (PP6)  
%shakura73, pacz78, lin87,

Accretion disks such as those surrounding black holes can develop both
MRI and GI \citep{menou01,goodman03}.  
Protoplanetary disks (PPDs) are also expected to be massive and  
magnetized in its earliest evolutionary phase \citep{inutsuka10}.  
The interplay between MRI and GI has been invoked to explain outbursts
in circumstellar disks \citep{armitage01, zhu10a, zhu10b,
  martin12b}, and predicts similar phenomenon in circumplanetary disks
\citep{lubow12}. This results from the development of `dead zones' 
--- magnetically inactive, laminar regions near the disk midplane --- with
magnetized layers above and below 
\citep{gammie96,martin12,landry13}. Mass accumulation in the dead zone can lead to
GI and trigger MRI through heating. In these models, the condition required
for MRI is realized through GI, but the MRI is unaffected by disk
self-gravity.

PPDs subject to both MRI and GI are often modeled through separate
turbulent viscosities in a hydrodynamical framework 
\citep{terquem08}. This implicitly assumes that the development of MRI 
and GI can be assessed independently. 
%That is, MRI is unaffected by
%disk self-gravity; and GI is unaffected by weak magnetization. 
Circumstellar disk models that explicitly combine the equations of magneto-hydrodynamics
and self-gravity have been limited to a few early simulations 
\citep{fromang04,fromang04b,fromang05}. It will be neccessary to %no
%major difference % kim/ostriker/stone
revisit and extend these pioneering calculations to fully explore the
impact of MRI and GI on the structure and evolution of PPDs. In
preparation of this, it is important to have a thorough understanding
of the stability properties of such systems.  
%non-ideal effects important

%why is GI and MRI independent 
%compressibility is not fundamental to MRI 
%horizontal wavenumber (gammie/balbus) 
%The effect of self-gravity on the linear MRI has not been considered, 
%since fluid compressibility is not fundamental to the MRI.

Since compressibility is not fundamental for the MRI, much of the
early stability calculations assume incompressible perturbations
\citep{goodman94,jin96}. However, recent works indicate
compressibility may be important under certain conditions,
such as strong fields \citep{kim00, pessah05,bonanno07}.   
Previous MRI studies have also focused on modes with vanishing
radial wavenumber, because they are the most unstable
\citep{sano99,reyes01}. Self-gravity has minimal effect on such
perturbations in a rotating disk. However, modes
with radial lengthscales on the order of the disk scale height may be
subject to self-gravity. It is of interest to generalize the MRI with
non-zero radial waveumbers to massive disks.

%Simiarly, studies of GI in accretion disks have mostly neglected
%magnetization \citep{mamat10,kim12}. 
%particularly in the context of

%Studies of GI in accretion disks do not often include magnetization
%\citep{mamat10,kim12}.

The effect of a magnetic field on the GI of rotating disks has been
considered by some authors.  
% \cite{lynden-bell66} found that a toroidal
%field can destabilize non-axisymmetric disturbances
Recently, \cite{lizano10} generalized the classic Toomre stability
criterion for razor-thin disks to include a vertical field. For
circumstellar disks, the authors concluded that the field is
stabilizing. This is consistent with previous analysis by
\cite{nakamura83} for three-dimensional (3D) uniformly rotating
disks. However, the GI of 3D differentially rotating disks have
mostly neglected magnetic fields \citep{mamat10,kim12}, but such disks
are subject to the MRI if magnetized. 
%how is the dispersion relation for GI affected by the presence of MRI?
%It is interesting to ask whether
%or not the presence of MRI affects GI. 
%
%It is desirable to clarify
%whether or not the presence of MRI modes affect the dispersion
%relation for GI.

%although MRI and GI are physically distinct 
%explicitly show

This paper marks the beginning of our study of magnetized,
self-gravitating PPDs. We start from linear calculations. 
Linear stability analyses have the advantage that a wide range of
parameters can be studied with negligible computational cost. This
allows one to indentify conditions, if any, under which MRI and GI
cannot be consdered independent. It is also important to have such 
calculations at hand to help benchmark and guide non-linear
simulations.

%This paper marks the beginning of our study of magnetized,
%self-gravitating PPDs and is organized as follows.

This paper is organized as follows. \S\ref{setup} lists the governing
equations and describes the disk equilibra under consideration. 
The linear problem is formulated in \S\ref{linear}. The impact
of self-gravity on the MRI is discussed in
\S\ref{result1}; gravitationally unstable disks are considered in
\S\ref{result2} and equilibra with an azimuthal field is explored in
\S\ref{result3}. We summarize results in \S\ref{summary}, including a
discussion of caveats and future directions.

\input{setup}
\input{linear}
\input{results}
\input{results_gi}
\input{results_azi} 
\input{summary}

\appendix
\input{appendix}

\bibliographystyle{apj}
\bibliography{ref}

\end{document}