### 2014-03-27 paper/results_gi.tex

\section{Gravitationally unstable disks} \label{result2}
Gravitational instability (GI) becomes possible in a sufficiently
massive and/or cold disk. We explore how the emergence of GI affects
the MRI, and vice versa. To do so, we compute unstbale modes for  
isothermal disks with $Q < 0.2 $ ($Q_\mathrm{2D}\lesssim 0.67$) which
permits gravitational instability as well as the MRI.

\subsection{Co-existence of MRI and GI}
We first consider ideal MHD by setting $\Lambda_0=100$ and $A=1$. 
Fig. \ref{compare_growth3} show growth rates for 
modes with $k_xH=1$ as a function of $\beta$ in disks with $Q=0.18$,
$Q=0.14$ and $Q=0.12$. The colorbar quantifies the importance of
self-gravity with 
\begin{align}
 \tau \equiv \frac{\avg{E_g}}{\avg{E_g} + \avg{E_m}}. 
\end{align} 
Thus, modes with $\tau  = 1$ are
energetically dominated by self-gravity (GI) and $\tau\ll1$ represents a mode
dominated by magnetic perturbations (MRI).

All three cases display distinct GI modes (red/brown branch). The GI
growth rates are $\gamma\simeq 0.25\Omega,\,0.6\Omega,\,0.8\Omega$ for
$Q=0.18,\,0.14,\,0.12$, respecitvely. GI is stabilized by magnetic
pressure for sufficiently small $\beta$, and the critical field
strength increases with the self-gravity \citep{nakamura83}. For
$Q=0.18$, GI is stabilized for $\beta \lesssim 15$. Nevertheless, 
the MRI branch for $\beta < 15$ becomes self-gravitating, so that
density perturbations are still expected even though  GI does not
formally operate.

%critical beta increases with $Q$

\begin{figure}
  \includegraphics[width=\linewidth,clip=true,trim=0cm 2.1cm 0cm
    0cm]{figures/compare_growth3_kx1_Q0d18.ps} 
  \includegraphics[width=\linewidth,clip=true,trim=0cm 2.1cm 0cm
    0.5cm]{figures/compare_growth3_kx1_Q0d14.ps} 
  \includegraphics[width=\linewidth,clip=true,trim=0cm 0cm 0cm
    0.5cm]{figures/compare_growth3_kx1_Q0d12.ps} 
  \caption{Growth rates for linear modes with $k_xH=1$ in an isothermal disk with
    $Q=0.18$ (top), $Q=0.14$ (middle) and $Q=0.14$ (bottom). The
    colorbar measures the importance of 
    self-gravity.  
    \label{compare_growth3}}
\end{figure}

The GI and MRI branches only interact when their growth rates
are similar. This is seen in Fig. \ref{compare_growth3} for $Q=0.18$
where the GI branch approaches an MRI branch at $\beta\simeq 25,\,
\gamma\simeq 0.2\Omega$. In fact following the red curve to smaller
$\beta$ indicates GI transitions to MRI. The
`gaps' in the GI and MRI branches for $Q=0.18$ and $Q=0.12$ may be due
to avoided crossing of modes, as seen in stars
\citep[e.g.][]{aizenman77} and accretion tori 
\citep[e.g.][]{christo93}. However, we cannot exclude the possibility 
that some modes may have been missed in the numerical search of
eigenfrequencies.

Thus, our results do not rigorously prove that the GI and MRI branches
do not intersect.  Nevertheless, the continous variation of $\tau$ with
respect to $\beta$ strongly suggest that unstable modes can transition
smoothly from magnetically-dominated to gravitationally-dominated at
low $\beta$.

\subsubsection{Case study} 
In reality, perturbations with a range of $k_x$ will be present for a
given set of disk parameters. Fig. \ref{compare_growth3_Q01d2} show
growth rates as a function of $k_x$ for an isothermal disk with
$Q=0.12,\,\beta=20$. MRI and GI are active with comparable growth
rates. All perturbations with $k_xH \lesssim 3.5$ grow dynamically
($\gamma\gtrsim 0.1\Omega$, or $\lesssim$ 1.6 orbits).

We also plot in Fig. \ref{compare_growth3_Q01d2} growth rates obtained 
from the Cowling approximation, which isolates MRI; and that from a
high-resistivity run, which isolates GI by allowing the field lines to
slip through the fluid. We refer to these as  
pure MRI and pure GI, respectively. 
For $k_xH\lesssim 0.2$, growth rates are equal to
those on the pure MRI and pure GI branches. That is, MRI and GI
operate independently until their growth rates become equal as a
function of $k_x$.

The dispersion relation $\gamma(k_x)$ increasingly deviates from the
pure GI/MRI curves with $k_x$, implying increasing interaction between
magnetic and density perturbations. 
Comparing pure GI (dashed line)  and the gravitationally-dominated
portions of $\gamma(k_x)$ indicate inclusion of magnetic field
stabilizes high-$k_x$ pure GI. (Note also the slight decrease in the most unstable
$k_x$.) Since axisymmetric GI is unaffected by shear, this
stabilization is due to magnetic pressure and is consistent with
\citep{lizano10}.

%3D geometry doesn't affect 
%axisymmetric GI doesn't care about shear 
%
%two legs of GI are not symmetric 
%MRI become self-gravitating
%GI with increasing magnetization

Comparing pure MRI (solid line) and the magnetically-dominated
portions of $\gamma(k_x)$ show that self-gravity can enhance or reduce
MRI growth rates depending on $k_x$, but the effect is small.

%magnetic pressure, small scale more effective 
%, reflected in the full range of $\tau$ displayed by the
%modes. 
%mode classification becomes difficult. 
%the point is: dispersion relation cannot be estimated by computing
%the MRI and GI separately

\begin{figure}
  \includegraphics[width=\linewidth]{figures/compare_growth3_Q0d12_B20.ps}  
  \caption{Growth rates of unstable modes in the massive isothermal
    disk with $Q=0.12$ and $\beta=20$, as a function of the horizontal
    wavenumber $k_x$. The colorbar measures the importance of
    self-gravity. The solid line corresponds to MRI modes in the
    Cowling approximation. The dashed line corresponds to pure GI
    modes, obtained by including a high resisitivity in the full problem. 
    \label{compare_growth3_Q01d2}}
\end{figure}

%\subsection{Layered disks}
%Can a disk with a dense, resistive midplane and a highly conducting
%atmosphere  be subject to both MRI and GI but at different heights?

%GI in the midplne and MRI in the surface?