### 2014-03-31 paper/results.tex

\section{MRI in self-gravitating disks}\label{result1}
In this section we focus on the MRI and use the vertical field setup
of case 1. We first consider MRI modes with negligble density/potential
perturbations to see the effect of self-gravity on the MRI through the
background stratification, then go on to examine MRI modes with
density/potential perturbations in massive disks.

\subsection{Influence of self-gravity on the MRI through the
  background equilibrium}
Here we use polytropic disks, so that the disk thickness is
well-defined. The upper disk boundary is set to 
$Z_s=0.99H$. We fix $\beta=100$ and $k_xH=0.1$ unless otherwise
stated.

%kx -> 0 so incompressible

\subsubsection{Uniform
  resistivity}   
%We calculate the MRI as a function of $Q$ with uniform
%resistivity ($A=1$) for mid-plane Elsasser numbers $\Lambda_0\in[0.3,10]$.    
Fig. \ref{compare_growth_poly_uniresis} plots MRI growth rates as a
function of $Q$ and $\Lambda_0$. The resistivity is uniform ($A=1$). 
%As expected growth rates generally decrease with decreasing 
%$\Lambda_0$ \citep{jin96}. 
In the limit of ideal MHD ($\Lambda_0>1$), there is negligible
dependence on $Q$. In the resistive limit 
($\Lambda_0<1$), however, growth rates decrease noticeably for $Q<0.5$
($Q_\mathrm{2D}\lesssim 1.5$). This shows 
show that disk self-gravity affects the MRI through the background
equilibrium, even though density and potential perturbations are  
unimportant (i.e. the linear response is non-self-gravitating).

\begin{figure}
  \includegraphics[width=\linewidth]{figures/compare_growth_poly_uniresis2}
  \caption{MRI growth rates as a function of $Q$ for mid-planee 
    Elsasser numbers $\Lambda_0=0.3$ (solid), $0.5$ (dotted), $1.0$
    (dashed) and $5.0$ (dash-dot).  
    \label{compare_growth_poly_uniresis}}
\end{figure}

\cite{sano99} found that for MRI to operate, its 
wavelength $\lambda$ should fit inside the disk. That is,   
\begin{align}\label{sano_crit}
  \lambda \equiv
  \mathrm{max}\left(\lambda_\mathrm{ideal},\lambda_\mathrm{resis}\right)\lesssim
  2H, 
\end{align}
where the MRI wavelengths are given by 
\begin{align}\label{lambda_ideal}
  \frac{\lambda_\mathrm{ideal}}{2H} = \frac{4\pi}{\sqrt{15}} f \hat{v}_A =
  \frac{4\pi f}{\sqrt{15\beta\hat{\rho}}}
\end{align}
for ideal MHD, and 
\begin{align}\label{lambda_resis}
  \frac{\lambda_\mathrm{resis}}{2H} = \frac{2\pi}{\sqrt{3}}\frac{\hat{\eta}}{\hat{v}_A f} =
  \frac{2\pi f}{\Lambda_0}\sqrt{\frac{\hat{\rho}}{3\beta}} 
\end{align}
in the limit of high resistivity.

Since $\hat{\rho}$ is weakly dependent
on $Q$ (Fig. \ref{eqm_den}), self-gravity only affects the
MRI through the factor $f$, which increases with decreasing $Q$ (see
Fig. \ref{plot_fq} in Appendix \ref{appen1}).   
This implies that sufficiently strong self-gravity can stabilize the
MRI by making $ 2H<\lambda $.  %these perts are not gravitationally
                               %unstable

In the ideal limit, we find $\lambda < 2H$ throughout most of the disk
for the values of $Q$ considered, so self-gravity does not affect
growth rates significantly. However, the wavelength of the
instability, in units of $H$, increases. This is shown in
Fig. \ref{compare_result_lambda10} which plots the magnetic energies
for $\Lambda_0=10$ and a range of $Q$ values. The number of vertical
nodes decrease with $Q$, i.e. the disk accommodates fewer wavelengths
because increasing vertical self-gravity makes it thinner.

%Weak variations in growth rates in the ideal limit, seen in 
%Fig. \ref{compare_growth_poly_uniresis}, were found to be associated
%with changes in the mode character: $\gamma$ decreases
%slightly when the mode begins to acquire an additional vertical node as $Q$ is
%increased.

%{\bf DISCUSS STRESS} Although for ideal MHD self-gravity makes no
%difference to the MRI growth rate, we find the associated stress does
%change with $Q$.

\begin{figure}
  \includegraphics[width=\linewidth]{figures/compare_result_lambda10}
  \caption{Magnetic energies as a function of height, in the limit of
    ideal MHD, for various strengths of self-gravity.   
    \label{compare_result_lambda10}}
\end{figure}

Self-gravity appreciably decreases the MRI growth rates in the
resistive limit. Fig. \ref{lambda_poly_resis} plots 
Eq. \ref{sano_crit} for $\Lambda_0=0.3$. In the non-self-gravitating
disk ($Q=4$) the instability criterion is marginally satisfied and the
MRI operates. As $Q$ decreases, 
Eq. \ref{sano_crit} is violated and the MRI growth rate is
significantly reduced. This is seen for $Q=0.2$ where $\lambda \geq 2H$ throughout
the disk. (The instability is not completely surpressed since
Eq. \ref{sano_crit} is only an approximation.) Although the function
$f(Q)$ does not change significantly for the range of $Q$ considered, the
dependence of $\lambda_\mathrm{resis}$ on $f(Q)$ is amplified by the
denominator $\Lambda_0<1$ in the resistive case. For resistive cases
we find no nodes in the magnetic energy $E_m$ except at end points,
i.e. only the longest wavelength survives.

\begin{figure}
  \includegraphics[width=\linewidth]{figures/lambda_poly_uniresis}
  \caption{Approximate wavelengths of the most unstable MRI modes as given by
    Eq. \ref{sano_crit}---\ref{lambda_resis}, normalized by the 
    disk thickness, as a function of height. MRI is expected to
    operate if $\lambda/2H\lesssim 1$. 
    \label{lambda_poly_resis}}
\end{figure}

\subsubsection{Layered 
  resistivity} 
Here we consider disks with mid-plane Elsasser number $\Lambda_0=0.1$
and a variable resistivity profile with
$A=10^2$. Fig. \ref{poly_layer} compares the magnetic  
energies for $Q=0.2,\,1$ and $4$. They have similar growth rates, $\gamma/\Omega
= 0.53,\,0.64$ and $0.66$, respectively. In the non-self-gravitating
limit ($Q=4$), the MRI is effectively surpressed for
$z\lesssim0.5H$. This is consistent with layered accretion. However,
in the massive disk ($Q=0.2$) the mode occupies a wider vertical
extent because its wavelength (in units of $H$) is larger. This
suggests that in massive disks the MRI is not well localized to the
active layer.

%MRI penetrates 
%d is not as well localized

\begin{figure}
  \includegraphics[width=\linewidth]{figures/compare_results_poly_layer_amp100}
  \caption{Magnetic energies as a function of height, for layered polytropic disks with
    such that the conductivity increases by a 
    factor of $10^2$ in going from the mid-plane to the upper disk
    boundary. The vertical lines indicate $\Lambda=1$ for each value
    of $Q$.
    \label{poly_layer}}
\end{figure}

\subsubsection{Dependence on $k_x$}
The above experiments show that with increasing 
disk self-gravity, the MRI becomes more global in the vertical
direction. We find a similar result in the horizontal direction. 
In Fig. \ref{compare_growth_poly_kx} we plot MRI growth rates as a
function of $k_x$ for a range of $Q$ values. Increasing self-gravity
decreases the cut-off horizontal wavenumber for the MRI. We checked
that these modes have negligible density perturbations. Then we can 
understand this result by invoking the instability criteria for  
incompressible MRI in an unstratified Keplerian disk,
\begin{align}
  v_A^2(k_z^2 + k_x^2) < 3\Omega^2,
\end{align}
where $k_z$ is a vertical wavenumber \citep{kim00}. Setting $k_z^2\sim
\Omega^2/v_A^2$ and non-dimensionalizing, we find
\begin{align} 
  k_xH \lesssim \frac{\sqrt{\beta}}{f},
\end{align}
where order-unity factors have been dropped. Despite a simplistic
approach, this demonstrates that with increasing self-gravity
(increasing $f$), we expect that MRI modes with small radial
lengthscales to be suppressed.

%, implying that
%small-scale MRI does not operate in massive disks.

\begin{figure}
  \includegraphics[width=\linewidth]{figures/compare_growth_poly_varQ_kx}
  \caption{MRI growth rates in self-gravitating polytropic disks, as a
    function of the horizontal wavenumber $k_x$. The disk is ideal
    ($\Lambda_0=10^2,\, A=1$) with $\beta = 40$. These modes have negligible
    density/potential perturbations.  
    \label{compare_growth_poly_kx}}
\end{figure}

%these modes are still non-sg
%use beta = 40 so only a small range of kx need to be simulated to
%demonstrate this effect
%{\bf cut-off $k_xH$ decreases with increasing SG -> small-scale MRI
%  stabilized by self-gravity}

\subsection{Influence of self-gravity on the MRI through the linear
  response}  
Our goal here is to examine whether or not self-gravity can amplify
the density perturbations caused by the MRI. We compute unstable modes
in a massive isothermal disk with $Q=0.2$ (corresponding to
$Q_\mathrm{2D}=0.72$), which is still expected to be marginally  
stable to gravitational instability \citep[][who find
a critical value of  $Q\simeq 0.2$]{mamat10}.   
The upper disk boundary is set to $Z_s=H$.

%resolution Nz=256

\subsubsection{Ideal disks}
We first consider ideal MHD by adopting a uniform 
resistivity with $\Lambda_0=100$. %Non-zero density perturbations
%require $k_x\neq0$ so we consider linear modes as a function of $k_x$.   
Fig. \ref{gravity_energy} plots MRI growth rates as a function of
$k_x$ for several values of $\beta$. The curves are color-coded
according to $\tau$. The potential perturbation is
negligible for all cases when $k_xH\lesssim 0.5$, since the MRI
becomes incompressible as $k_x\to 0$.

\begin{figure}
  \includegraphics[width=\linewidth]{figures/compare_energy_ideal}
  \caption{Growth rates of MRI modes in an isothermal self-gravitating
    disk with $Q=0.2$ ($Q_\mathrm{2D}=0.72$) in the limit of ideal MHD
    ($\Lambda_0=10^2$), for a range of field strengths $\beta$. The
    colorbar measures the importance of self-gravity by $\tau$.  
    %{\bf ADD BETA=8}
    \label{gravity_energy}}
\end{figure}

For $\beta\gg 1$, i.e. a weak field, density perturbations are
negligible and the standard MRI operates. However, as $\beta$ is
lowered and the MRI growth rate reduced, we find non-negligible
potential perturbation for $k_xH=O(1)$. This suggets that in a
strongly magnetized disk that still permits the MRI, the associated
density perturbation can be important when the disk is
self-gravitating. %but formall stable

%critical beta, below which MRI is surpressed, increases with the
%strength of self-gravity.

%Fig. \ref{mri_massive} shows an example of a MRI mode with
%non-negligible density perturbation. The growth rate $\gamma\simeq
%0.4\Omega$ is comparable to that of the most unstable mode
%($\gamma\simeq 0.6\Omega$ for $k_x\simeq0$). 
%We therefore expect significant density perturbations to grow
%dynamically, even though the disk is expected to be gravitationally
%stable.

%\begin{figure}
%  \includegraphics[width=\linewidth]{figures/result_ideal_sg}
%  \caption{Example of a linear MRI mode with non-negligible density and
%    potential perturbation, in a strongly magnetized massive disk
%    ($\beta=10$ and $Q=0.2$) in the limit of ideal MHD 
%    ($\Lambda_0=10^2$).   
%    \label{mri_massive}}
%\end{figure}

\subsubsection{Resistive disks}
We repeat the above calculation for resistive disks, but fix
$\beta=100$ and vary the mid-plane Elasser number $\Lambda_0$. Growth
rates are shown in Fig. \ref{gravity_energy_resis}. 
Interestingly, in the highly resistive case
the magnetic and gravitational energies can be comparable. For 
$\Lambda_0=0.1$ and $k_xH\simeq1.3$, $\tau\sim 0.3$ which corresponds
to $\avg{E_g}\sim 0.5\avg{E_m}$.  
Fig. \ref{mri_massive_cowling}, we compare the magnetic
energy of this mode to that computed in in the \emph{Cowling  
  approximation}, where the Poisson equation is ignored in the 
linearized equations and the potential perturbation set to zero  
(formally letting $Q\to\infty$ in Eq. \ref{lin_poisson}).  
The growth rate increases slightly when self-gravity is included in the linear
response, since self-gravity is usually destabilizing. However, the
similarity between $\gamma$ and $E_m$ indicates that the instability
in the self-gravitating calculation is fundamentally the MRI.

\begin{figure}
  \includegraphics[width=\linewidth]{figures/compare_energy_resis}
  \caption{
    Growth rates of MRI modes in an isothermal self-gravitating
    disk with $Q=0.2$ ($Q_\mathrm{2D}=0.72$) at fixed $\beta=100$, 
    for a range of mid-plane Elasser numbers. The resistivity is
    uniform. The colorbar measures the importance of self-gravity by $\tau$. 
    \label{gravity_energy_resis}}
\end{figure}

\begin{figure}
  \includegraphics[width=\linewidth]{figures/compare_result_cowling}
  \caption{Magnetic energy associated with the linear mode with
    largest gravitational-to-magnetic energy ratio in
    Fig. \ref{gravity_energy_resis} (solid) compared with that computed
    under the Cowling approximation (dotted). %{\bf REDO FIGURE}
    \label{mri_massive_cowling}}
\end{figure}

Fig. \ref{mri_massive_resis} plots the energies assocaited with the 
MRI mode discussed above. 
%with $k_xH\sim1.3$ in the $\Lambda_0=0.1$  
%disk. 
The gravitational energy exceeds the 
magnetic energy near the midplane ($z\lesssim0.2H$). The growth rate $\gamma=0.25\Omega$
is not much smaller than that of the most unstable mode
($\gamma=0.36\Omega$ for $k_xH=0.1$), so significant density
perturbations are expected to grow on dynamical timescales for this
system, even though GI should not develop.

\begin{figure}
  \includegraphics[width=\linewidth]{figures/result_resis_sg}
  \caption{Example of a resistive MRI mode with significant
    gravitational potential perturbation.  The vertical line
    inidicates $\Lambda=1$.  
    \label{mri_massive_resis}}
\end{figure}

\subsubsection{Qualitative interpretation} 
To make sense of the above results, we first return to ideal MHD and 
consider regions close to the disk midplane ($z\sim 0$), where
self-gravity is expected to be most important. For this discussion we
will ignore stratification and set $d^2/dz^2\to -k_z^2$. The governing
equations are then

\begin{align}
  &  0= v_A^2k^2\dvx + \imgi\sigma\left(\imgi\sigma \dvx - 2\Omega\dvy + \imgi k_x \w\right),\label{simp1}\\
  &  0= v_A^2k_z^2\left(\dvy + \frac{\imgi S
  }{\sigma}\dvx\right) + \imgi\sigma\left(\imgi\sigma\dvy +
  \frac{\kappa^2}{2\Omega}\dvx\right),\label{simp2}\\
  & 0 = -k_z^2\w + \frac{\sigma^2}{c_s^2}W + \sigma k_x \dvx, \label{simp3}\\
  & 0 = k^2\dphi + \frac{\Omega^2}{c_s^2Q}W,\label{simp4}
\end{align}
where $k^2 = k_z^2 + k_x^2$. We imagine an iterative procedure to
solve the above equations, starting from the Cowling approximation. 
%defined to be the
%solution to Eq. \ref{simp1}---\ref{simp3} with potential perturbation
%set to zero. 
This is the standard MRI and we denote the solution as
$\dvx^{(0)}$, $\dvy^{(0)}$ and $W^{(0)}$.

We now include self-gravity. In general, the compressible MRI has
$W^{(0)}\neq0$. The Poisson equation implies
an associated potential perturbation,    
\begin{align} 
  \dphi = -\frac{\Omega^2}{c_s^2 Q k^2} W^{(0)}.
\end{align}
Physically, we expect $k^2\geq0$, so that a local positive (negative) density
perturbation causes a negative (positive) local potential
perturbation. We then insert $\dphi$ back into the momentum and
continuity equations, and ask how does this potential perturbation
modify the Cowling solution? Writing $\dvx^{(0)} \to \dvx^{(0)} +
\dvx^{(1)}$ and similarly for $\dvy $ and $ W$, we find 
%regard $\dphi$ to alter the Cowling solution, so
%that 
%Then 
\begin{align}
  &   k_x\sigma \dphi = v_A^2k^2\dvx^{(1)} + \imgi\sigma\left[\imgi\sigma
  \dvx^{(1)} - 2\Omega\dvy^{(1)} + \imgi k_x W^{(1)}\right], \label{simp_pert1}\\ 
  &  0= v_A^2k_z^2\left[\dvy^{(1)} + \frac{\imgi S
    }{\sigma}\dvx^{(1)}\right] + \imgi\sigma\left[\imgi\sigma\dvy^{(1)} +
  \frac{\kappa^2}{2\Omega}\dvx^{(1)}\right],\label{simp_pert2}\\
  & k_z^2\dphi  = \left(\frac{\sigma^2}{c_s^2}-k_z^2\right)W^{(1)} +
  \sigma k_x \dvx^{(1)}.\label{simp_pert3} 
\end{align}
Now, if the perturbations to the magnetic field remain
unchanged, i.e. the mode remains close to the standard MRI, then
$\dvx^{(1)} \sim 0$ and $\dvy^{(1)}\sim0$, so Eq. \ref{simp_pert2} is
satisfied. Eq. \ref{simp_pert1} then require $\dphi +
W^{(1)}\sim0$. This is compatible with Eq. \ref{simp_pert3} if 
\begin{align}
  \left|k_z^2\right| \gg \left|\frac{\sigma^2}{c_s^2}\right|. \label{cond}
\end{align}
For the ideal MRI,  we assume $k_z\sim \Omega/v_A$. Then
$|\sigma^2/c_s^2k_z^2|\sim |\sigma^2/\Omega^2\beta|\ll1$ because
$|\sigma|\lesssim \Omega$ and we are considering $\beta\gtrsim 10$. 
Thus Eq. \ref{cond} is generally satisfied.

The above assumptions imply
\begin{align}
  W^{(1)} \sim \frac{\Omega^2}{c_s^2 Q k^2} W^{(0)},\label{feedback}
\end{align} 
which indicates a non-zero density perturbation due to the
MRI can be amplified by self-gravity. Now, for $k_xH\sim 1$ we have
$|k_z^2/k_x^2|\sim \beta/f\gg1$ because $f=O(1)$ and $\beta\gtrsim10$
for the cases considered above. Then
\begin{align}
  \left|\frac{W^{(1)}}{W^{(0)}}\right| \sim \frac{1}{Q\beta}, 
\end{align}
 suggesting stronger amplification of the density field by 
 self-gravity with increasing field strength (decreasing
 $\beta$).

% not expect significant potential perturbations 
%Conversely, for strong fields, only a single MRI wavelength occupies
%the fluid column.

 The above arguments can be adapted to the resistive
 limit. Eq. \ref{simp3}---\ref{simp4} are unchanged, while resistive
 terms appearing in Eq. \ref{simp1}---\ref{simp2} only involve the
 potential perturbation through $\w$. For the resistive MRI we assume
 $k_z\sim v_A/\eta$ and $|\sigma|\sim v_A^2/\eta = \Lambda\Omega$
 \citep{sano99}. Then $|\sigma^2/c_s^2k_z^2|\sim 1/\beta \ll 1$ so
 Eq. \ref{cond} is satisfied. Noting that $k_z^2\sim
 \Lambda^2\Omega^2\beta/c_s^2$, the feedback equation becomes
 \begin{align}
   \left|\frac{W^{(1)}}{W^{(0)}}\right| \sim
   \frac{1}{Q\left(f^2\hat{k}_x^2 + \beta\Lambda^2\right)},
 \end{align}
 so increasing the resistivity (decreasing $\Lambda$) should enhance
 density perturbations.

 For weak fields, the MRI has small vertical wavelength,
 implying rapid density variations with height. The magnitude of the
 corresponding gravitational potential is then expected to be small
 because of internal cancellations, since the potential is essentially
 the integral of the density field.