### 2014-05-29 paper/apj/response.txt
final revised version

I thank the reviewer for comments and questions. I have performed
additional calculations and extended the discussion section to address
the issues raised below. Revisions to the manuscript are made in
bold.

**********************************************************************

1) For a fixed value of B_z, massive disks would have larger values of
plasma beta, compared to the less massive disks. In actual disks
around young stars or compact objects, how large is the required value
of B_z in order for self-gravity to affect MRI modes?  
**********************************************************************
I have expanded Section 7 to include a quantitative discussion on the
limiting field strength to permit the MRI. I show that a larger plasma
beta is required in strongly self-gravitating disks. I also estimate
the upper limit to the field strength, B_z, for Q<<1.

In addition, Fig. 2 in Section 4.1.1 has been updated to include a new
calculation, \beta=25 and \Lambda_0=10. This case demonstrates the
stabilizing effect of self-gravity in the ideal limit. 
**********************************************************************

2) The amplitude of the resistivity variation from the disk surface to
the midplane (A=eta(0) / eta(infinity) ) is fixed to 100. If you take
larger values of A, such as A=10^3 or 10^4, as seen in Figure 1 of
Gressel et al. (2012), is there a qualitative change in the result?
For example, in the Q=0.2 case in Figure 5, does the magnetic energy
still penetrate into the resistive zone? In a massive disk, the value
of A is expected to be large. Could you comment on the qualitative
tendency for high A?  
**********************************************************************
Additional calculations with A=10^3 and A=10^4 are described in
Section 4.1.2. Similar results were obtained for large A with small
Lambda_0. When the midplane resistivity (or Lambda_0) is fixed and A
is increased, the magnetic energy in the resistive zone lessens. For
A=10^3 or A=10^4 and Lambda_0=0.1, both self-gravitating and
non-self-gravitating disks show little magnetic energy in the
resistive zones. However, lowering Lambda_0 with these larger values
of A yields similar results to the case in Fig. 5.

**********************************************************************
3) Section 3.1.1 Case 1: vertical field
Please mention that the odd modes are considered in Section 6, and
cite references for explaining how the boundary conditions, Equation
(29), are derived, e.g., Sano and Miyama (1999). 
**********************************************************************
Section 3.1 has been re-structured. I describe where in the results
sections are the different field setups considered. I also make the
reference to Sano & Miyama (1999).

**********************************************************************
4) Caption of Figure 11
Q=0.14 (bottom) --> Q=0.12 (bottom) 
**********************************************************************
Corrected.
**********************************************************************