### 2014-06-03 paper/cover.tex

% Cover letter using letter.sty
\documentclass{letter} % Uses 10pt
%Use \documentstyle[newcent]{letter} for New Century Schoolbook postscript font
% the following commands control the margins:
\topmargin=-1in    % Make letterhead start about 1 inch from top of page 
\textheight=4in  % text height can be bigger for a longer letter
\oddsidemargin=0pt % leftmargin is 1 inch
\textwidth=6.5in   % textwidth of 6.5in leaves 1 inch for right margin

\begin{document}

\signature{Min-Kai Lin}           % name for signature 
\longindentation=0pt                       % needed to get closing flush left
\let\raggedleft\raggedright                % needed to get date flush left

\begin{letter}{The Editor, The Astrophysical Journal}

\begin{flushleft}
{\large\bf Min-Kai Lin}
\end{flushleft}
\medskip\hrule height 1pt
\begin{flushright}
\hfill Canadian Institute for Theoretical Astrophysics, \\
\hfill 60 St. George Street, Toronto, ON \\
\hfill M5S 3H8, Canada\\
\hfill +1 647 896 4314
\end{flushright} 
\vfill % forces letterhead to top of page

\opening{Dear Editor,}

I am writing to submit the revised manuscript `Linear stability of
magnetized massive protoplanetary disks', in response to the recent
reviewer's report. Revisions are made in bold, and I attach my
response to the reviewer.

\closing{Yours sincerely,}

%\encl{}  				% Enclosures

\end{letter}

\end{document}